---
title: "The ChIPanalyser User's Guide"
author: "Patrick Martin"
output:
    pdf_document: null
    keep_tex: yes
    number_sections: no
    html_document: default
    code_folding: show
package: '`ChIPanalyser'''
vignette: |
    %\VignetteIndexEntry{ChIPanalyser User's Guide for Genetic Algorithms}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
---

# Introduction - What is this package?
ChIPanalyser provides a quick an easy method to predict and explain TF binding.
The package uses a statistical thermodynamic framework to model the binding of
proteins to DNA.

The model assumes that there are four main driver to TF binding:

* Chromatin State affinity
* Binding Energy
* Number of bound Molecules
* Scaling Factor modulating binding affinity

While binding energy is given by PWM matrices, the other parameters will be
inferred by maximizing or minimizing a goodness of fit score between ChIP
data and ChIPanalyser predictions. To do so, ChIPanalyser provides a genetic
algorithm to infer optimal values for each parameter.

# Methods - The Chromatin State Model

The chromatin state model is described by the following equation derived from statistical thermodynamics:

$$P(N,a,\lambda,\omega)_j = \frac{N \cdot a_{j} \cdot e^{(\frac{1}{\lambda} \cdot \omega_{j})}}{N \cdot a_{j} \cdot e^{(\frac{1}{\lambda} \cdot \omega_{j})}+ L \cdot n \cdot [a_{i} \cdot e^{(\frac{1}{\lambda} \cdot \omega_{j})}]_ {i}} $$


* $N$ , the average number of bound molecules
* $a_{j}$ , chromatin state affinity at site j
* $\omega$ , the binding energy required for a TF to bind to site $j$ - in the form of a Position Weight Matrix Score
* $\lambda$ , a scaling factor for the Position Weight Matrix score
* $L$ , the length of the genome of interest
* $n$ , the ploidy level of the organism  


Chromatin state affinity is defined as the following:

$$a_{j} = \sum_{k} \alpha_k \cdot c_{j}^k$$

with $\alpha$ the chromatin state affinity scores for a given TF and $c$ the chromatin state at site $j$.

Shortly, the model above generates ChIP like profiles and describes the affinity of a TF to chromatin states. 
The affinity scores $a$, the number of bound molecules $N$, and lambda $\lambda$ are inferred by optimising the fit between 
predicted profiles and ChIP data using a genetic algortihm. 

# Using ChIPanalyser 

## Loading data
To demonstrate the use of ChIPanalyser and the genetic algorithm, we provide some internal data sets. We complement this data with 
external data source such a DNA sequence sets taken from the `BSgenome`package suit. 

### ChIPanalyser 
First, we load ChIPanlyser and the internal data 
```{r loading_ChIPanalyser, eval = TRUE, echo = TRUE}
library(ChIPanalyser)
# Input data 
data(ChIPanalyserData)
# PFM Matrix 
PFM <- file.path(system.file("extdata",package="ChIPanalyser"),"BEAF-32.pfm")

```

### External Data 
To reduce the size of the internal data size, we complement our data with the _Drosophila_ DNA sequence taken from the BSgenome package suit. 
```{r bsgenome, eval = TRUE, echo = TRUE}
library(BSgenome.Dmelanogaster.UCSC.dm6)

DNASequenceSet <- getSeq(BSgenome.Dmelanogaster.UCSC.dm6)

```

### Input data : What is it? 
Now we that have loaded some data, what exactly are we looking at? The environment should contain the following new objects:

* *chip* - GRanges object containing ChIP scores. 
Essentially, ChIP data that will be used to _train_ our model. 

* *cs* - GRanges object containing Chromatin State (CS) information. 
Essentially, where each CS can be found in the genome. 

* *top* - GRanges object containing regions of interest. 
Esssentially, the genomic regions we will use for training and testing. 

* *PFM* - Path to file containing the Position Frequency Matrix. 
Essentially, this represents the binding affinity of our TF (here we are using BEAF-32)

* *DNASequenceSet* - DNAStringSet containing the DNASequenceSet of the organism we are looking at. 

## Setting Parameters
The next step consitst in setting initial parameters to run the genetic algorithm. 
```{r GA_params, eval = TRUE, echo = TRUE}
# Number of individuals per generation 
pop <- 10 

# Number of generations 
gen <- 10

# Mutation Probability 
mut <- 0.3

# Children - Number of ofspring passed to the next generation 
child <- 2 

# Method - Goodness of fit metric used to optimise the Genetic algorithm 
method <- "MSE"

```

We can also define which parameters we wish to optimise. In this example, we will optimise the number of bound molecules ($N$),
lambda ($\lambda$), the PWM Threshold and 11 different chromatin states. 
```{r opti_params, eval = TRUE, echo = TRUE}
# Parameters to optimised
params <- c("N","lambda","PWMThreshold", paste0("CS",seq(1:11)))
```

Alternatively, we can also set custom ranges for the parameters. 
```{r opti_params2, class.source = 'fold-hide',eval = FALSE,echo= TRUE}
params_custom <- vector("list", 14)
names(params_custom) <- c("N","lambda","PWMThreshold", paste0("CS",seq(1:11)))

# vector in the format of min value, max value and number of values 
params_custom$N <- c(1,100000,20)

params_custom$lambda <- c(0.5,5,10)

# Bound between 0 and 1 
params_custom$PWMThreshold <- c(0,1,10)

# Bound between 0 and 1 
CS <- c(0,1,5)
CS_loc <- grep("CS",names(params_custom))
for(i in CS_loc){
    params_custom[[i]] <- CS
}
```

## Initializing ChIPanalyser
### Building Initial objects 
The first step of any ChIPanalyser anaylsis is to create parameter object. These object contain input paramters and hold basic data. 
Here, we will load the PFM as PWM and compute the Base Pair Frequency from the DNASequenceSet. 

```{r gpp, eval = TRUE, echo = TRUE}
GPP <- genomicProfiles(PFM=PFM,PFMFormat="JASPAR", BPFrequency=DNASequenceSet)

```

### Generating a starting population 
For clairty, we will show what a starting population looks like. This step is not stricly required as the `evolve` function can take the 
number of individuals in the poppulation (defined above by `pop`) and the parameters to be optimised (defined above by `params`).
```{r generateStartingPopulation, class.source = 'fold-hide', eval = TRUE, echo = TRUE}
start_pop <- generateStartingPopulation(pop, params)
```

### Pre-processing ChIP data
We will pre-proccess the ChIP data by reducing noise and converting GRanges to a ChIPscore object. This object contains normalised and smoothed
ChIP scores that will be used to train and test the model. 

```{r preprocess, eval = TRUE, echo = TRUE}
chipProfile <- processingChIP(chip,loci = top)

# Splitting data into training and testing
# We recommend setting dist to 20/80. However, here we only have 2 loci. 
splitdata <- splitData(chipProfile,dist = c(50,50))

trainingSet <- splitdata$trainingSet
testingSet <- splitdata$testingSet
```
    
## Evolution 




    
